### 2015-10-15 Schloss_Cluster_PeerJ_2015.Rmd
Edit Abstract and Introduction

---
title: "***De novo* clustering methods out-perform reference-based methods for assigning 16S rRNA gene sequences to operational taxonomic units**"
author: Patrick D. Schloss* and Sarah L. Westcott
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
bibliography: references.bib
output:
  pdf_document:
    includes:
      in_header: header.tex
csl: peerj.csl
fontsize: 11pt
geometry: margin=1.0in
---

## To Do:
* references
* revise
* add sensitivity/specificity of closed-reference searching

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)

inline_hook <- function(x){
  if(is.numeric(x)){
    paste(format(x,big.mark=',', digits=2, scientific=FALSE))
  }
}
knitr::knit_hooks$set(inline=inline_hook)
```

* To whom correspondence should be addressed.
pschloss@umich.edu

Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI

\newpage
## Abstract

**Background.** 16S rRNA gene sequences are routinely assigned to operational taxonomic units (OTUs) that are then used to analyze complex microbial communities. A number of methods have been employed to carry out the assignment of 16S rRNA gene sequences to OTUs leading to confusion over which method is the best. A recent study suggested that a clustering method should be selected based on its ability to generate stable OTU assignments that do not change as additional sequences are added to the dataset. In contrast, we contend that the ability of the method to properly represent the distances between the sequences is more important.

**Methods.** Our analysis implemented five *de novo* clustering algorithms including the single linkage, complete linkage, average linkage, abundance-based greedy clustering, distance-based greedy clustering and two reference-based methods including the open and closed-reference methods. By varying the number of sequences sampled from each of two previously published datasets we used the Matthew’s Correlation Coefficient (MCC) to assess the quality and stability of OTU assignments.

**Results.** The stability of OTU assignments did not reflect the quality of the assignments. Depending on the dataset being analyzed, the average linkage and the distance and abundance-based greedy clustering methods generated more robust OTUs than the open and closed-reference methods. Further interrogation of the reference-based methods indicated that when USEARCH is used to identify the closest reference, the OTU assignments were sensitive to the order of the reference sequences. When USEARCH was replaced by VSEARCH the quality and stability of the OTU assignments by the reference-based methods were improved. Finally, we demonstrated that for the greedy algorithms VSEARCH produced assignments that were comparable to those produced by USEARCH making VSEARCH a viable free and open source alternative to USEARCH.

**Discussion.** This study demonstrates that the quality of clustering assignments needs to be assessed for multiple methods per dataset to identify the optimal clustering method for the dataset. The analysis also calls into question the quality and stability of OTU assignments generated by the open and closed-reference methods as implemented in current versions of QIIME.

\newpage

## Introduction

The ability to affordably generate millions of 16S rRNA gene sequences has allowed microbial ecologists to thoroughly characterize the microbial community composition of hundreds of samples. To simplify the complexity of these large datasets, it is helpful to cluster sequences into meaningful bins. These bins, commonly known as operational taxonomic units (OTUs), are  used to compare the biodiversity contained within and between different samples [@schloss2011assessing]. Such comparisons have enabled researchers to characterize the microbiota associated with the human body [e.g. @human2012structure], soil [e.g. @shade2013streptomycin], aquatic ecosystems [e.g. @gilbert2012defining], and numerous other environments. Within the field of microbial ecology, a convention has emerged where sequences are clustered into OTUs using a threshold of 97% similarity or a distance of 3%. One advantage of the OTU-based approach is that the definition of the bins is operational and can be changed to suit the needs of the particular project. However, with the dissemination of clustering methods within software such as mothur [@schloss2009introducing], QIIME [@Caporaso2010], and other tools [@edgar2010search, @sun2009esprit, @10.7717/peerj.593], it is important to understand how different clustering methods implement this conventional OTU threshold. Furthermore, it is necessary to understand how method choice affects the precision and accuracy of assigning sequences to OTUs. Broadly speaking, three approaches have been developed to assign sequences to OTUs.

The first approach has been referred to as phylotyping [@schloss2011assessing] or closed reference clustering [@navas2013advancing]. This approach involves comparing sequences to a curated database and then clustering sequences into the same OTU that are similar to the same reference sequence. Reference-based clustering methods suffer when the reference does not adequately reflect the biodiversity of the community. If a large fraction of sequences are novel, then they cannot be assigned to an OTU. In addition, the reference sequences used in this application are selected because they are less than 97% similar to each other over the full length of the gene; however, it is known that the commonly used variable regions within the 16S rRNA gene do not evolve at the same rate as the full-length gene. Thus, a sequence representing a fragment of the gene may be more than 97% similar to multiple reference sequences. Therefore, defining OTUs in the closed-reference approach is complicated because although two sequences might be 97% similar to the same reference sequence, they may only be 94% similar to each other. A subtle alternative to this approach is to use a classifier to assign a taxonomy to each sequence so that sequences can be clustered at a desired level within the Linnean taxonomic hierarchy [@schloss2011assessing]. The strength of the reference based approach is that the methods are generally fast, scaling linearly with the number of sequences being clustered.

The second approach has been referred to as distance-based [@schloss2011assessing] or *de novo* clustering [@navas2013advancing]. In this method, the distance between sequences is used to cluster sequences into OTUs rather than the distance to a reference database. In contrast to the efficiency of closed-reference clustering, the speed of hierarchical *de novo* clustering methods scale quadratically with the number of unique sequences. The expansion in sequencing throughput combined with sequencing errors inflates the number of unique sequences resulting in the need for large amounts of memory and time to cluster the sequences. If error rates can be reduced through stringent quality control measures then these problems can be overcome [@Kozich2013]. As an alternative, heuristics have been developed to approximate the clustering of hierarchical methods [@edgar2010search, @sun2009esprit, @10.7717/peerj.593]. One critique of *de novo* approaches is that OTU assignments are sensitive to the input order of the sequences [REFS]. Whether the differences in assignments is meaningful is unclear; however, the variation in results could represent equally valid clustering of the data. The strength of *de novo* clustering is its independence of references for carrying out the clustering step. After clustering, the classification of each sequence can be used to obtain a consensus classification for the OTU [@Schloss2011]. For this reason, *de novo* clustering has been preferred across the field.

The third approach, open-reference clustering, is a hybrid of the closed-reference and *de novo* approaches [@Navas2013]. Open-reference clustering method involves performing closed-reference clustering followed by *de novo* clustering on those sequences that are not sufficiently similar to the reference. In theory, this method should exploit the strengths of both closed-referene and *de novo* clustering; however, the different OTU definitions employed by both approaches poses a possible problem when the methods are combined. An alternative to this approach has been to classify sequences to a bacterial family or genus and then assigned to OTUs within those levels using the average linkage method [@Schloss2011]. For example, all sequences assigned to the *Porphyromonadaceae* would then be assigned to OTUs using the average linkage method using a 3% distance threshold. Those sequences that did not classify to a known family would also be clustered using the average linkage method. An advantage of this approach is that it lends itself nicely to parallelization since each taxonomic group is seen as being independent and can be processed separately. Such an approach would overcome the difficulty of mixing OTU definitions between the closed-reference and *de novo* approaches.

The growth in options for assigning sequences using each of these three broad approaches has been considerable. It has been difficult to objectively assess the quality of OTU assignments. Some have focused on the time and memory required to process a dataset (REFS). These are valid parameters to assess when judging a clustering method, but have little to say about the quality of the clustering. Others have attempted to judge the quality of an method by its ability to generate data that parallels classification data (REFS). This approach is problematic because bacterial taxonomy often reflects historical biases amongst bacterial systematicists. Furthermore, it is well known that the rates of evolution across lineages are not the same (REFS). We recently proposed an approach for evaluating OTU assignments using the distance between sequences (REF). Those sequences that were similar to each other and found in the same OTU were called true positives while those that were similar and found in different OTUs were called false negatives. Meanwhile, those sequences that were different from each other and found in the same OTU were called false positives and those that were dissimilar and found in different OTUs were called true negatives. Counting the frequency of these different classes allowed us to judge how each method balanced the ratio of true positives and negatives to false positives and negatives using the Matthew's correlation coefficient (MCC; REF). This is an objective approach to assessing the quality of the OTU assignments.

A recent analysis by He and colleagues (REF) attempted to characterize the three general clustering approaches by focusing on what they called stability. They defined stability as the ability of an method to provide the same clustering on a subset of the data as was found in the full dataset. Their concept of stability did not account for the accuracy of the OTU assignments and instead focused on the precision of the assignments. A method may be very precise, but low in accuracy. In the current analysis, we assessed the accuracy and precision of the various clustering methods. Building on our previous analysis of clustering methods, our hypothesis was that the methods praised by the He study for their stability actually suffered a lack of accuracy. In addition, we assess these parameters in light of sequence quality using the original 454 dataset and a larger and more modern dataset generated using the MiSeq platform.

## Results and Discussion
```{r he_stability_analysis}
he_stability <- read.table(file="data/process/he.mcc_ref.summary", header=T)
```

***Summary and replication of He study.*** We sought to identify the more critical analyses performed in the He study. Similar to their study, we obtained the Canadian soil dataset from Roesch et al. (REF) and processed the sequences as described in their analysis. In our opinion there were three important tests.

First, we sought to quantify whether the clustering observed for a subset of the data represented the same clustering that was found with the full dataset. The He study found that when they used the open and closed-reference methods clusters formed using the subset data most closely resembled those of the full dataset. Among the *de novo* clustering methods they observed that the abundance-based greedy clustering (AGC) method followed by single linkage (SL), distance-based greedy clustering (DGC), complete linkage (CL), and average linkage (AL) methods. They observed a broad range of MCC values among their AL replicates, which was not explained by the authors. We first sought to assess calculated the MCC for for each of the clustering methods using 20, 40, 60, and 80% relative to the clusters formed by the methods using the full dataset (Figure 1A). Because a random number generator is used in some of methods to break ties where pairs of sequences have the same distance between them, similar to the He study, we replicated each method and subsample 30 times. Across these sequencing depths, we observed that the stability of the SL and CL methods were highly sensitive to sampling effort relative to the AL, AGC, and DGC methods (Figure 1A). Our results (Figure 1B) largely confirmed those of Figure 4C of He study with one notable exception. In the He study, when comparing the OTUs formed using the AL method with 60% of the data, they observed a mean MCC value of approximately 0.63 (95% confidence interval between approximately 0.15 and 0.75). In contrast, we observed a mean value of `r round(he_stability[he_stability$method=="an" & he_stability$fraction==0.6, "mean"], 2)` (95% confidence interval between `r round(he_stability[he_stability$method=="an" & he_stability$fraction==0.6, "lci"], 2)` and `r round(he_stability[he_stability$method=="an" & he_stability$fraction==0.6, "uci"], 2)`). This result suggests that the AL method was far more stabile than indicated in the He study. Regardless, it supports the assertion that the clustering observed for the subset of the data does not represent the same clustering that was found with the full dataset; however, the significance of these differences is unclear.

Second, rarefaction curves calculated using clusters obtained using a portion of the dataset did not overlap with rarefaction curves generated using clusters generated from the full dataset. This result was originally observed in the Roesch study using the complete linkage method and was reproduced in the He study where this result was more pronounced problem when using the CL, SL, and DGC methods relative to the other methods. The He and Roesch studies both found that the CL method produced fewer OTUs in the subset than in the rarefied data. Expanding the analysis to other methods, the He study found that the SL method produced more OTUs, the AGC produced fewer, and the other methods produced similar numbers of OTUs than expected when comparing the subsetted data to the rarefied data. Our results support those of these previous studies (Figure 2). It was clear that inter-method differences were generally more pronounced than the differences observed between rarefying from the full dataset and from clustering the subsetted data. The number of OTUs observed was largest using the CL method, followed by the open-reference method. The AL, AGC, and DGC methods all provided comparable numbers of OTUs. Finally, the closed-referene and SL methods generated the fewest number of OTUs.

Third, the authors attempted to describe the effects of the clustering instability on comparisons of communities. They used Adonis to test whether the community structure represented in subsetted communities resembled that of the full dataset. Inspection of these results indicates that although they were able to detect significant p-values, they were were of marginal biological significance. Adonis R statistics close to zero indicate the community structures from the full and subsetted datasets overlapped while values of one indicate the communities are completely different. The He study observed adonis R statistics of 0.02 (closed-reference), 0.03 (open-reference), 0.07 (CL, AGC, DGC), and 0.16 (SL and AL). Regardless of the statistical or biological significance of these results, the analysis does not make sense since the *de novo* and open-reference approaches do not consistently label the OTUs that sequences belong to when the clustering methods are run multiple times with different random number seeds. To overcome this, the authors selected representative sequences from each OTU and used those representative sequences to link OTU assignments between the different sized sequence sets. The justification for this analysis is specious as the OTU assignments are based on the data available in the dataset when the sequences are clustered and comparing assignments in this manner are irreconcilable. It is not surprising that the only analysis that did not provide a significant p-value was for the closed-reference analysis, which is the only analysis that provides consistent OTU labels. Because this analysis was so poorly designed, we did not seek to reproduce it.

***Methods vary in their accuracy.*** More important than the stability of OTUs is whether sequences are assigned to the correct OTUs. A method can generate highly stable OTUs, but the OTUs may be meaningless by poorly representing a specified cutoff in the assignment of sequences to those OTUs. To assess the accuracy and precision of the various methods, we made use of the pairwise distance between the unique sequences to count the number of true positives and negatives and the number of false positives and negatives for each method and sampling depth so that we could calculate the average MCC value as a measure of a method's accuracy and its variation as a measure of its precision. We made three important observations. First, each of the *de novo* methods varied in how sensitive their MCC values were to additional sequences (Figure 1C). The SL and CL methods were the most sensitive; however, the MCC values using 80 and 100% of the data did not vary meaningfully when using *de novo* methods. Second, the AL method out-performed the other methods followed by DGC, AGC, CL, open-reference, and closed-reference, and SL (Figure 1D). Third, with the possible exception of the CL method, the MCC values for each of the only demonstrated a small amount of variation between runs of the method with a different ordering of the input sequences. This indicates that although there may be variation between executions of the same method, they produce OTU assignments that are equally good. Revisiting the concept of stability, we question the value of obtaining stable OTUs when the the full dataset is not optimally assigned to OTUs. Our analysis indicates that the best method for assigning the Canadian soils sequences to OTUs using a 97% threshold is the AL method.

```{r region_overlap}
n_overlap <- read.table(file="data/gg_13_8/gg_13_8_97.overlap.count", row.names=1)
n_v4_overlap <- n_overlap["v4", 1]
n_v9_overlap <- n_overlap["v9", 1]
n_v19_total <- n_overlap["v19", 1]
```

```{r miseq_analysis}
miseq_info <- read.table(file="data/miseq/miseq.seq.info", row.names=1)
n_groups <- miseq_info["n_groups",1]
n_miseq_seqs <- miseq_info["n_miseq_seqs",1]
n_unique_miseq_seqs <- miseq_info["n_unique_miseq_seqs",1]

miseq_stability <- read.table(file="data/process/miseq.mcc_ref.summary", header=T)
miseq_mcc <- read.table(file="data/process/miseq.mcc.summary", header=T)
```

***Deep sampling of 16S rRNA genes.*** Three factors make the Canadian soil dataset less than desirable to evaluate clustering methods. First, the Canadian soil dataset that the He study used to analyze the stability of OTUs is one of the earliest 16S rRNA gene sequence datasets published using the 454 FLX platform. Developments in sequencing technology now permits the sequencing of millions of sequences for a study. In addition, because the original phred quality scores and flowgram data are not available, it was not possible to adequately remove sequencing errors (REFS). The large number of sequences that one would expect to remain in the dataset are likely to negatively affect the performance of all of the clustering methods. Second, the dataset used in the He study covered the V9 region of the 16S rRNA gene. For a variety of reasons, this region is not well represented in databases, including the reference database used by the closed and open-reference methods. Of the `r n_v19_total` sequences in the default QIIME database, only `r n_v4_overlap` fully cover the V9 region. In contrast, `r n_v4_overlap` of the sequences fully covered the V4 region. Inadequate coverage of the V9 region would adversely affect the ability of the reference-based methods to assign sequences to OTUs. Third, our previous analysis has shown that the V9 region evolves at a rate much slower than the rest of the gene. With these points in mind, we compared the clustering assignment for each of these methods using a time series experiment that was obtained using mouse stool (REFS). The MiSeq platform was used to generate `r n_miseq_seqs` sequences from the V4 region of the 16S rRNA gene of `r n_groups` samples. Parallel sequencing of a mock community indicated that the sequencing error rate was approximately 0.02% (REFS). Although no dataset is perfect for exhaustively testing these clustering methods, this dataset was useful for demonstrating several points. First, when using 60% of the data the stability relationships amongst the different methods were similar to what we observed using the He dataset (Figure 3AB). With the exception for the clusters generated using CL, the methods all performed very well with stabilities greater than `r round(min(miseq_stability[miseq_stability$fraction=="0.6" & miseq_stability$method != "fn","mean"]), 2)`. Second, the MCC values calculated relative to the distances between sequences were generally higher than was observed for the He dataset for all of the methods except for CL and SL. Surprisingly, the MCC values for the DGC (`r round(miseq_mcc[miseq_mcc$fraction==1.0 & miseq_mcc$method == "dgc","mean"],2)`)  and AGC (`r round(miseq_mcc[miseq_mcc$fraction==1.0 & miseq_mcc$method == "agc","mean"],2)`) methods were comparable to the AL method (`r round(miseq_mcc[miseq_mcc$fraction==1.0 & miseq_mcc$method == "an","mean"],2)`; Figure 3CD). This result suggests that the optimal method may be database-dependent. Finally, as was observed with the He dataset, there was little variation in MCC values observed among the 30 randomizations. Therefore, although the methods have a stochastic component, the OTU assignments do not vary meaningfully. The results from both the Canadian soil and murine microbiota datasets demonstrate that the *de novo* methods can generate very stable OTU assignments, as defined by the He study, and that the assignments are highly reproducible. Most importantly, these analyses demonstrate that the OTU assignments using the AL, AGC, and DGC *de novo* methods are consistently more robust than either of the reference-based methods.

***Evalution of an open-source alternative to USEARCH.*** The AGC and DGC methods appear to perform as well and possibly than the hierarchical clustering methods for some datasets. These methods utilize the USEARCH program (REF), which is not available as open source code and is only available for free to academic users as a 32-bit program. Access for non-academic users and those needing the 64-bit version is available commercially from the developer. An alternative to USEARCH is VSEARCH, which is being developed in parallel to USEARCH as a open-source alternative. One subtle difference between the two programs is that USEARCH employs a heuristic to determine when to stop searching a database because it think it has found the best hit or does not believe that the database contains a hit. This means that the best possible match is not necessarily returned once a highly similar hit is obtained. In contrast, VSEARCH does not utilize the heuristic, which the VSEARCH developers claim enhances the sensitivity relative to USEARCH. Using the two datasets, we determined whether the AGC and DGC methods as implemented by the two methods yielded sequencing assignments of similar quality. In general the overall trends that we observed with the USEARCH-version of AGC and DGC were also observed with the VSEARCH-verion of the methods. When we compared the two implementations of the AGC and DGC methods, the OTUs generated by the VSEARCH-verion of the methods were as stable or more stable than the USEARCH-version when using 60% of the datasets. In addition, the MCC values for the entire datasets, calculated relative to the distance matrix, were virtually indistinguishable (Figure 4). These results are a strong indication that VSEARCH is a suitable and possibly better replacement for USEARCH for executing the AGC and DGC methods.

```{r rand_ref_analysis}
rand_ref_u <- read.table(file="data/rand_ref/hits.uclosed.summary", row.names=1)
rand_ref_v <- read.table(file="data/rand_ref/hits.vclosed.summary", row.names=1)
```

***Problems with reference-based clustering in general and as implemented in QIIME.*** The He study indicated that the closed-reference method generated perfectly stable OTUs. This was unsurprising since, by definition, the method represents a one-to-one mapping of reads to a reference and treats the input sequences independently. An important test that was not performed in the He study was to determine whether the clustering was sensitive to the order of the sequences in the database. The default QIIME database, which was used in the He study and by others, contains full-length sequences that are at most 97% similar to each other. We randomized the order of the reference sequences 30 times and used them to carry out the closed-reference method with the full MiSeq dataset, which contained `r n_unique_miseq_seqs` unique sequences. Surprisingly, we observed that the number of OTUs generated was not the same in each of the randomizations. On average there were `r round(rand_ref_u["mean", 1], 2)` sequences that mapped to a reference OTU per randomization (range from `r round(rand_ref_u["min", 1], 2)` to `r round(rand_ref_u["max", 1], 2)`). The original ordering of the reference resulted in `r rand_ref_u["original", 1]` sequences being mapped, less than the minimum observed number of mapped sequences when the references were randomized. This surprising result was likely due to the performance of the USEARCH heuristic. As with the AGC and DGC methods, if the order of the reference sequences differed between runs, then the heuristic criteria would be triggered at different points. To test this further, we substituted VSEARCH for USEARCH in the closed-reference method. When we used VSEARCH the original ordering of the reference sequences and all randomizations were able to map `r rand_ref_v["original", 1]` sequences to reference OTUs. That VSEARCH mapped more sequences than even the maximum number of sequences found across the USEARCH randomizations supports the VSEARCH developers' assertion that it has greater sensitivity than USEARCH.

```{r database_analysis}
small_dists <- read.table(file="data/gg_13_8/gg_13_8_97.small_dist.count", row.names=1)
n_small_v4_dist <- small_dists["v4",1]
n_small_v19_dist <- small_dists["v19",1]

n_v4_unique <- length(scan(file="data/gg_13_8/gg_13_8_97.v4_ref.names", what="", sep="\n", quiet=T))

u_overlap <- read.table(file="data/rand_ref/overlap.uclosed.summary")
v_overlap <- read.table(file="data/rand_ref/overlap.vclosed.summary")

dups <- read.table(file="data/gg_13_8/duplicate.analysis")
max_taxa <- dups[dups$V3 == max(dups$V3),]
```

We also observed that regardless of whether we used USEARCH or VSEARCH, the reference OTU labels that were assigned to each OTU differed between randomizations. When we used USEARCH to perform closed-reference clustering, an average of `r round(100*u_overlap["orig", "mean"], 1)`% of the labels were shared between pairs of the 30 randomizations (range=`r round(100*u_overlap["orig", "min"], 1)` to `r round(100*u_overlap["orig", "max"], 1)`). If we instead used VSEARCH an average of `r round(100*v_overlap["orig", "mean"], 1)`% of the labels were shared between pairs of the 30 randomizations (range=`r round(100*v_overlap["orig", "min"], 1)` to `r round(100*v_overlap["orig", "max"], 1)`). To better understand this result, we further analyzed the reference database that is used in QIIME. We hypothesized that within a given region there would be sequences that were more than 97% similar and possibly identical to each other. When a sequence was used to search the randomized databases, it would encounter a different reference sequence as the first match with each randomization. Among those reference sequences that fully overlap the V4 region, there were `r n_small_v19_dist` pairs of sequences that were more than 97% similar to each other over the full length of the 16S rRNA gene. When the extracted V4 sequences were dereplicated, we identified `r n_v4_unique` unique sequences. Among these dereplicated V4 sequences there were `r n_small_v4_dist` pairs of sequences that were more than 97% similar to each other. The presence of duplicate V4 reference sequences explains the lack of labeling stability when using either USEARCH or VSEARCH to carry out the closed-reference method. We suspect that the reference database was designed to only include sequences that were at most 97% similar to each other way to overcome the limitations of the USEARCH search heuristic.

Beyond comparing the abundance of specific OTUs across samples, the reference database is used in the open and closed-reference methods to generate OTU labels that are used in several downstream applications. It is commonly used to extract information from a reference phylogenetic tree to carrying out UniFrac-based analyses (REFS) and to identify reference genomes for performing analyses such as PICRUSt (REFS). Because these downstream applications depend on the correct and unique labeling of the OTUs, the lack of stability of the labeling is problematic. As one illustration of the effects that incorrect labels would have on an analysis, we asked whether the duplicate sequences had the same taxonomies. Among the `r sum(dups$V2==2)` reference sequences that had one duplicate, `r sum(dups$V2==2 & dups$V3==2)` had discordant taxonomies. Furthermore, among those `r sum(dups$V2 > 2)` sequences with two or more duplicates, `r sum(dups$V2>2 & dups$V3>=2)` had discordant taxonomies. Two sequences mapped to `r max_taxa[1,2]-1` and `r max_taxa[2,2]-1` duplicate sequences and both contained `r max_taxa[1,3]` different taxonomies. There was also a sequence had `r max(dups[,2])-1` duplicates and contained `r dups[which.max(dups[,2]),3]` different taxonomies. Together, these results demonstrate some of the considerable problems with the reference-based clustering of sequences.

## Conclusions
It is worth noting that the entire design of the He study was artificial. First, their analysis was based on a single soil sample. Researchers generally have dozens or hundreds of samples that are pooled and clustered together to enable comparison across samples. Second, all of the sequence data from these datasets is pooled for a single analysis. It is unclear why anyone would ever perform an analysis based on a subset of their data (REFS). Because of these points, the value of identifying stable OTUs is unclear. Greater emphasis should be placed on obtaining an optimal balance between splitting similar sequences into separate OTUs and merging disparate sequences into the same OTU. Through the use of the pairwise distances between sequences, we were able to use the MCC to demonstrate that, in general, the AL, AGC, and DGC methods perform better than the others. Although there is concern that running the methods multiple times yields different clusterings, we have shown that there is little variation in their MCC values. This suggests that the different clusterings by the same method are equally good. Finally, it is impossible to obtain a clustering with no false positives or false negatives and the optimal method may vary by dataset. With this in mind, researchers are encouraged to run the AL and VSEARCH DGC methods and calculate and report their MCC values.

Our analysis of those methods that implemented USEARCH as a method for clustering sequences revealed that its heuristic limited its sensitivity. When we replaced USEARCH with VSEARCH, the clustering quality improved. Although there may be parameters in USEARCH that can be tuned to improve the heuristic, these parameters are likely dataset dependent. Based on the data presented in this study, its availability as an open source, and free program, VSEARCH should replace USEARCH in these clustering methods. Furthermore, although not tested in our study, VSEARCH can be parallelized leading to potentially significant improvements in speed. Although USEARCh and VSEARCH do utilize aligned sequences, it is important to note that a sequence curation pipeline including denoising, alignment, trimming to a consistent region of the 16S rRNA gene, and chimera checking are critical to making proper inferences (REF).

For the first time, we assessed the ability of reference-based clustering methods to capture the actual distance between the sequences in a dataset. Several studies have lauded both the open and closed-reference approaches for generating reproducible clusterings, yet we showed that both reference-based approaches did a poor job of representing the distance between the sequences compared to the *de novo* approaches. So, although the clusterings are reproducible and stable across a range of library sizes, the clusterings are a poor representation of the data. We also observed that the clusterings were not actually reproducible when the order of the reference sequences was randomize. When USEARCH was used, the actual number of sequences that mapped to the reference changed depending on the order of the reference. Perhaps most alarming is that the default order of the database provided the worst sensitivity of any of the randomizations we attempted. Even when we used VSEARCH to perform closed-reference clustering and were able to obtain a consistent clusterings, we observed that the labels on the OTUs differed between randomizations. Because the OTU labels are frequently used to identify representative sequences for those OTUs, variation in labels, often representing different taxonomic groups, will have a detrimental effect on the interpretation of downstream analyses.

Because the open-reference method is a hybrid of the closed-reference and DGC methods, it is also negatively affected by the problems with using USEARCH for both methods. An added problem with the open-reference method is that the two phases of the method employ different thresholds. In the closed-reference step, sequences must be within a threshold of a reference to be in the same OTU. This means that two sequences that are 97% similar to a reference and are joined into the same OTU, may only be 94% similar to each other. In the DGC step, the goal is to approximate the AL method which requires that, on average, the sequences within an OTU are at least 97% similar to each other. The end result of the open-reference approach is that sequences that are similar to previously observed sequences are clustered with one threshold while those that are not similar to previously observed sequences are clustered with a different threshold.

As the throughput of sequencing technologies have improved, method development to keep pace. *De novo* clustering methods are considerably slower and more computationally intensive than reference-based methods and the greedy *de novo* methods are faster than the hierarchical. In our experience (REFS), the most significant detriment to execution speed of the *de novo* methods has been in adequate removal of sequencing error. As the rate of sequencing error increases so do the number of unique sequences that must be clustered. The speed of the *de novo* methods scales approximately quadratically, so that doubling the number of sequences results in a four-fold increase in the time required to execute the method. The rapid expansion in sequencing throughput has been likened to the Red Queen in Lewis Carroll's, *Through the Looking-Glass* who must run in place to keep up to her changing surroundings (REF). Microbial ecologists must continue to refine clustering methods to better handle the size of the datasets, but they must also take steps to improve the quality of the underlying data. Ultimately, objective standards must be applied to assess the quality of the data and the quality of OTU clustering.

## Methods

***454 FLX-generated Roesch Canadian soil dataset***
After obtaining the 16S rRNA gene fragments from GenBank (accessions EF308591-EF361836), we followed the methods outlined by the He study by removing any sequence that contained an ambiguous base, was identified as being a chimera, and fell outside a defined sequence length. Although they reported observing a total of 50,542 sequences that were represented by 13,293 unique sequences, we obtained a total of `r length(scan(file="data/he/canada_soil.good.unique.pick.redundant.fasta", what="", sep="\n", quiet=T))/2` sequences that were represented by `r length(scan(file="data/he/canada_soil.good.unique.pick.fasta", what="", sep="\n", quiet=T))/2` unique sequences. Similar to the He study, we randomly sampled, without replacement, 20, 40, 60, and 80% of the sequences from the full data set. The random sampling was repeated 30 times. The order of the sequences in the full dataset was randomly permuted without replacement to generate an additional 30 datasets. To perform the hierarchical clustering methods and to generate a distance matrix we followed their approach of the He study by calculating distances based on pairwise global alignments using the `pairwise.dist` command in mothur using the default Needleman-Wunsch alignment method and parameters. It should be noted that this method has been strongly discouraged (REFS). Execution of the hierarchical clustering methods was performed as described in the original He study using mothur (v.1.37) and using the QIIME (v.1.9.1) parameter profiles provided in the supplementary material from the He study for the greedy and reference-based clustering methods.

***MiSeq-generated Murine gut microbiota dataset***
The murine 16S rRNA gene sequence data generated from the V4 region using an Illumina MiSeq was obtained from http:/www.mothur.org/MiSeqDevelopmentData/StabilityNoMetaG.tar and was processed as outlined in the original study (REF). Briefly, 250 nt long read pairs were assembled into contigs by aligning the reads and correcting discordant base calls by requiring one of the base calls to have a Phred quality score at least 6 points higher than the other. Sequences where it was not possible to resolve the disagreement were culled from the dataset. The sequences were then aligned to a SILVA reference alignment (REF) and any reads that aligned outside of the V4 region were removed from the dataset. Sequences were pre-clustered by combining the abundances of sequences that were within 2 nt of a more abundant sequence. Each of the samples was then screened for chimeric sequences using the default parameters in UCHIME (REF). The resulting sequences were processed in the same manner as the Canadian soil dataset with the exception that the distance matrices were calculated based on the SILVA-based alignment.

***Analysis of reference database***
We utilized the 97% OTUs greengenes reference sequence and taxonomy data (v.13.8) that accompanies the QIIME installation. Because the greengenes reference alignment does a poor job of representing the secondary structure of the 16S rRNA gene (REF), we realigned the fasta sequences to a SILVA reference alignment to identify the V4 region of the sequences.

***Calculation of Matthew's Correlation Coefficient (MCC)***
The MCC was calculated by two approaches in this study. In both methods we used only the dereplicated sequence lists. First, we calculated the MCC to determine the stability of OTU assignments following the approach of the He study. We assumed that the clusters obtained from the 30 randomized full datasets were correct. We counted the number of sequence pairs that were in the same OTU for the subsetted dataset and the full dataset (i.e. true positives; TP), that were in different OTUs for the subsetted dataset and the full dataset (i.e. true negatives; TN), that were in the same OTU for the subsetted dataset and different OTUs in the full dataset (i.e. false positives; FP), and that were in different OTUs for the subsetted dataset and the same OTU in the full dataset (i.e. false negatives; FN). For each set of 30 random subsamplings of the dataset, we counted these parameters against the 30 randomizations of the full dataset. This gave 900 comparisons for each fraction of sequences being used in the analysis. The Matthew's correlation coefficient was then calculated as:

$$
MCC = \frac{TP \times TN - FP \times FN}{\sqrt{(TP +FP)(TP+FN)(TN+FP)(TN+FN)} }
$$

Second, we calculated MCC to determine the quality of the clusterings as previously described (REF). Briefly, we compared the OTU assignments for pairs of sequences to the distance matrix that was calculated between all pairs of aligned sequences. For each dataset that was clustered, those sequences that were in the same OTU and had a distance less than 3% were TPs, those that were in different OTUs and had a distance greater than 3% were TNs, those that were in the same OTU and had a distance greater than 3% were FPs, and those that were in different OTUs and had a distance less than 3% were FNs. The MCC was counted for each dataset using the formula above as implemented in the `sens.spec` command in mothur.

***Software availability***
A reproducible workflow including all scripts and this manuscript as a literate programming document are available at https://github.com/SchlossLab/Schloss_Cluster_PeerJ_2015. The workflow utilizes QIIME (v.1.9.1; REFS), mothur (v.1.37.0; REFS), USEARCH (v.1.5.0; REFS), VSEARCH (v.1.5.0; https://github.com/torognes/vsearch), and R (v.3.2; REFS).

\newpage

## Figures

**Figure 1. Comparison of the stability (A, B) and quality (C, D) of *de novo* and reference-based clustering methods using the Canadian soil dataset.** The average stability of the OTUs were determined by calculating the MCC with respect to the OTU assignments for the full dataset using varying sized subsamples (A). Thirty randomizations were performed for each fraction of the dataset and the average and 95% confidence interval are presented when using 60% of the data. The quality of the OTUs were determined by calculating the MCC with respect to the distances between the sequences using varying sized subsamples (C). Thirty randomizations were performed for each fraction of the dataset and the average and 95% confidence interval are presented when using the full dataset (D). The vertical gray line indicates in A and C indicates the fraction of the dataset represented in B and D, respectively.

**Figure 2. The clustering methods varied in their ability to generate the same number of OTUs using a subset of the data as were observed when the full dataset was rarefied.** The subsetted data are depicted by closed circles and the data from the rarefied full dataset is depicted by the open circles.

**Figure 3. Comparison of the stability (A, B) and quality (C, D) of *de novo* and reference-based clustering methods using the murine dataset.** The average stability of the OTUs were determined by calculating the MCC with respect to the OTU assignments for the full dataset using varying sized subsamples (A). Thirty randomizations were performed for each fraction of the dataset and the average and 95% confidence interval are presented when using 60% of the data. The quality of the OTUs were determined by calculating the MCC with respect to the distances between the sequences using varying sized subsamples (C). Thirty randomizations were performed for each fraction of the dataset and the average and 95% confidence interval are presented when using the full dataset (D). The vertical gray line indicates in A and C indicates the fraction of the dataset represented in B and D, respectively.

**Figure 4. The VSEARCH OTUs generated by the AGC and DGC methods were comparable to those generated using USEARCH.**

\newpage

## References